Geometric Filters for Metric Spaces

### 2018-11-29 paper/draft.tex

\Function{ApproximateNearestNeighbor}{$P, q, \eps$}
    \State {$[p_1, \dots, p_n] \gets \mbox{random permutation of }P$}
    \State { $a_i \gets 0$}
    \Comment {$a_i$ is lower bound for $\dist(p_i, q)$}
    \State {$ c \gets p_1, \quad v \gets \dist(p_1, q)$}
    \Comment {$c$ keeps the current candidate}
    \State \Call {UpdateBounds}{$p_1, v$}
    \For{$i = 2\dots n$}
        \If {$a_i \geq \frac{v}{1+\eps}$}
            \State {\textbf{continue}}
        \Else
            \State {Compute $r_i = \dist(p_i, q)$}
            \State \Call{UpdateBounds}{$p_i, r_i$}
            \If {$r_i < v$}
                \State {$c \gets p_i,\quad v\gets r_i$}
            \EndIf
        \EndIf
    \EndFor
    \State \Return {$c, v$}
\EndFunction
\end{algorithmic}

>>>>

The algorithm that we analyze can be roughly described as follows. 
Pick a random point $p_i$,
compute $\dist(p_i, q)$ and use this distance combined with pairwise distances $\dist(p_i, p_j)$
to update lower bounds for other distances $\dist(p_j, q)$.
If we see that the lower bound for $\dist(p_j, q)$ is larger than $\dist(p_i, q) / (1+\eps)$, then 
we can discard the point $p_j$. If all points were discarded, we found an approximate nearest neighbor;
otherwise pick a random point $p_j$ from the remaining ones and compute $\dist(p_j, q)$.
There are two possibilities: if $\dist(p_j, q) < \dist(p_i, q) / (1+\eps)$,
then we choose $p_j$ as the current candidate for an approximate nearest neighbor,
otherwise we keep $p_i$ in this role. In any case, the new computed distance
may also allow us to discard more than one point, as figure \ref{fig:discarded_balls} shows. In
this figure points inside the shaded smaller ball cannot be a better candidate,
and the same is true for the points in the shaded area outside of the larger ball.
The precise radii of these balls are given in the following lemma.
